---
title: "R Notebook"
output: html_notebook
---
---
title: "Scottish Health Boards"
output: html_notebook
---

> Copyright Scottish Government, contains Ordnance Survey data © Crown copyright and database right (insert year); The dataset covers all of Scotland, and is coterminous with Local Authorities (OS BoundaryLine, October 2017 with amendments at Keltybridge and Westfield); Open Government Licence (<http://www.nationalarchives.gov.uk/doc/open-government-licence/>)

Created from a [tutorial](https://github.com/mmoench16/Personal-Website/blob/b2fca6b7bda4d0f4b21eb30c1338ad199979633f/content/post/2018-09-04-shapefiles-in-r-with-ggplot2-rgdal.Rmd) by mmoench16 with an up to date data file

```{r}
library(tidyverse)
library(rgdal)
library(here)
library(broom)
```

```{r, eval = FALSE}
NHSBoards <- readOGR(dsn = "SG_NHS_HealthBoards_2019.shp", 
                     layer = "SG_NHS_HealthBoards_2019")
```

```{r, eval = FALSE}
# tidy() comes from the package library(broom)
# This produces a simple outline map of the NHS Boards
NHSBoards_tidy <- tidy(NHSBoards)
ggplot(NHSBoards_tidy, aes(x = long, y = lat, group = group)) +
  geom_polygon(color = "black", size = 0.1, fill = "seagreen") +
  coord_equal() +
  theme_minimal()
```

```{r, eval=FALSE}
NHSBoards$id <- row.names(NHSBoards)
NHSBoards_tidy <- left_join(NHSBoards_tidy, NHSBoards@data)
```

