---
title: "Covid Hospital Activity Analysis"
output:
  html_document:
    df_print: paged
---

```{r}
# How does Covid affect hospital activity?
# Was there a change post Q1 2020?
```


```{r, echo=FALSE, message=FALSE, results='hide'}
library(tidyverse)
library(sf)
library(janitor)
library(jsonlite)

library(rgdal)
library(broom)
require(readr)
require(here)
library(slider)
library(lubridate)
library(infer)
```
```{r}
# Dataset required
# Hospital Location
df_hospital_location <- read_csv("clean_data/df_hospital_location.csv")
df_cov_hb_ans <- read_csv("clean_data/df_cov_hb_ans.csv")
```

```{r}


# covid_hospital_activity_age & sex dataset
cov_hb_ans <- df_cov_hb_ans %>% 
  select(-percent_variation) %>% 
  filter(admission_type != "All") %>% 
  mutate(week = week(week_ending),
         year = as.character(year))

# For all trend, maybe a slider
# ALL GENERAL TREND, SEX = ALL, AGE = ALL
all_trend_hb_ans <- cov_hb_ans %>% 
  filter(age_group == "All ages" & sex == "All") %>% 
      group_by(year, week_ending, admission_type) %>% 
  summarise(number_admissions = sum(number_admissions)/1000,
            average20182019  = sum(average20182019)/1000) %>% 
    mutate(
    moving_avg_2021 = slide_dbl(
      .x = number_admissions,
      .f = ~mean(.,na.rm = TRUE),
# 2 weeks moving average
      .before = 2
    )  ) %>% 
          mutate(
    moving_avg_1819 = slide_dbl(
      .x = average20182019,
      .f = ~mean(.,na.rm = TRUE),
# 2 weeks moving average = 7
      .before = 2
    )  ) %>%
  # Find the winter quarter
  mutate(quarter = quarter(week_ending),
         quarter = case_when(
           quarter == "1" ~ 1,
           quarter == "2" ~ 3,
           quarter == "3" ~ 3,
           quarter == "4" ~ 3
         ))

# Plot ALL TREND
all_trend_hb_ans %>% 
  ggplot()+
  geom_line(aes(x = week_ending, y = moving_avg_2021, colour = admission_type), size=1.0)+
  geom_line(aes(x = week_ending, y = moving_avg_1819, colour = admission_type), linetype = "dashed")+
  geom_rect(aes(xmin = week_ending, xmax = dplyr::lead(week_ending), ymin = -Inf, ymax = Inf, fill = factor(quarter)), alpha = .3, show.legend = TRUE) +
    scale_fill_manual(labels = c("Winter", "Other"), values = alpha(c("darkblue","transparent"), .1))+
  
    geom_vline(xintercept = as.Date(ymd("2020-03-16")), linetype= 2, color = "black", size=0.6)+
  
  annotate(geom = "text",
             label = c("Lockdown"),
             x = c( as.Date(ymd("2020-03-16"))),
             y = c(26),
             vjust = 1,
           hjust = -.05) +
  
  theme(panel.background = element_rect(fill = 'transparent'),
        panel.grid.minor = element_line(colour = 'lightgrey'),
        panel.grid.major = element_line(colour = 'lightgrey'),
                 legend.position="top")+
    scale_x_date(date_breaks = "4 months",
                 date_minor_breaks = "1 months",
                 date_labels = "%b-%y") +
    labs(title = "2 weeks Moving Average Hospital Admission",
       subtitle = "solid = 2020-2021 data | dash = 2018-2019 data",
    colour = "Admission Type",
    fill = "Seasons"
  )+
  xlab("Time (date)")+
  ylab("Average Admission Count (Per 1'000)")



# When input isnt ALL Ages
  # geom_area(aes(x = week_ending, y = moving_avg_2021, fill = quarter == 4))+
  # geom_rect(aes(NULL, NULL, xmin = start, xmax = end, fill = quarter), ymin = yrng[1], ymax = yrng[2], data = all_trend_hb_ans)+
  # scale_fill_manual(values = alpha(c("blue", "red","red" ,"red"), .3))+
```

```{r}
age_group_trend_hb_ans <- cov_hb_ans %>% 
  filter(sex == "All" & age_group != "All ages") %>% 
  mutate(age_group = factor(age_group, levels = c("Under 5","5 - 14" ,"15 - 44", "45 - 64", "65 - 74", "75 - 84", "85 and over")
  )) %>%
  arrange(age_group) %>% 
      group_by(year, week_ending, age_group) %>% 
  summarise(number_admissions = sum(number_admissions)/1000) %>% 
    mutate(
    moving_avg_2021 = slide_dbl(
      .x = number_admissions,
      .f = ~mean(.,na.rm = TRUE),
# 2 weeks moving average
      .before = 2
    )  ) %>% 
  mutate(quarter = quarter(week_ending))


cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

age_group_trend_hb_ans %>% 
  ggplot()+
    geom_line(aes(x = week_ending, y = moving_avg_2021, group = age_group, colour = age_group))+
  
    geom_rect(aes(xmin = week_ending, xmax = dplyr::lead(week_ending), ymin = -Inf, ymax = Inf, fill = factor(quarter)), alpha = .3, show.legend = FALSE) +
    scale_fill_manual(values = alpha(c("darkblue", "transparent" ,"transparent" ,"transparent"), .4))+
      geom_vline(xintercept = as.Date(ymd("2020-03-16")), linetype= 2, color = "black", size=0.6)+
  annotate(geom = "text",
             label = c("Lockdown"),
             x = c( as.Date(ymd("2020-03-16"))),
             y = c(6.1),
             vjust = 1,
           hjust = -.05) +
  labs(title = "2 weeks Moving Average Hospital Admission Per Age Group in 2020 - 2021")+
  xlab("Time (date)")+
  ylab("Average Admission Count (Per 1'000)")+
  scale_colour_brewer(palette="Set1")+ 
    theme(panel.background = element_rect(fill = 'transparent'),
        panel.grid.minor = element_line(colour = 'lightgrey'),
        panel.grid.major = element_line(colour = 'lightgrey'),
                 legend.position="top")+
    scale_x_date(date_breaks = "4 months",
                 date_minor_breaks = "1 months",
                 date_labels = "%b-%y")

```

```{r}
# Display General Stat : ALL 
df_general_stat <- read_csv("clean_data/df_day_trend_loc.csv")

general_stat <- df_general_stat %>% 
  select(date, ca, ca_name, daily_positive, cumulative_positive, daily_deaths, cumulative_deaths) %>% 
  arrange(date) %>% 
  group_by(ca) %>% 
  slice_tail(n  = 1)

all_general_stat <- general_stat %>% 
  ungroup() %>% 
  group_by(date) %>% 
  summarise(daily_positive = sum(daily_positive),
            cumulative_positive = sum(cumulative_positive),
            daily_deaths = sum(daily_deaths),
            cumulative_deaths = sum(cumulative_deaths))


#specific stat per local authority
specific_general_stat <- general_stat %>% 
  ungroup() %>% 
  group_by(date, ca_name) %>% 
  filter (ca_name == "Midlothian") %>%  # FILTER BY LOCAL AUTHORITY
  summarise(daily_positive = sum(daily_positive),
            cumulative_positive = sum(cumulative_positive),
            daily_deaths = sum(daily_deaths),
            cumulative_deaths = sum(cumulative_deaths))

#Local Authority list
ca_name_list = factor(general_stat$ca_name)

all_general_stat

specific_general_stat
ca_name_list
```
# hypothesis testing
```{r}

hypo_all_trend_hb_ans <- all_trend_hb_ans %>% 
  select(year, week_ending, admission_type, number_admissions, average20182019) %>% 
  mutate(quarter = quarter(week_ending)) %>% 
  relocate(quarter, .after = year) %>% 
  rename(cov_adm = number_admissions,
         pre_cov_adm = average20182019) %>% 
    ungroup() %>% 
    mutate(cov_adm = cov_adm*1000,
           pre_cov_adm = pre_cov_adm*1000) %>%  
  mutate(cov_adm_rate = c(0, diff(cov_adm)),
         pre_cov_adm_rate = c(0, diff(pre_cov_adm)))

# Histogram for hypothesis test
hypo_all_trend_hb_ans %>% 
  ggplot()+
  geom_histogram(aes(x = cov_adm_rate), alpha = 0.5, fill = "red", bin = 30, col = 'white')+
    geom_histogram(aes(x = pre_cov_adm_rate), alpha = 0.5, fill = "blue", bin = 30, col = 'white')


hypo_all_trend_hb_ans %>% 
  ggplot()+
  geom_line(aes(x = week_ending, y = cov_adm_rate), colour = 'red')+
  geom_line(aes(x = week_ending, y = pre_cov_adm_rate), colour = 'blue')





hypo_all_trend_hb_ans %>% 
  ggplot()+
  geom_boxplot(aes(x = cov_adm_rate))
  
hypo_all_trend_hb_ans %>% 
  ggplot()+
  geom_boxplot(aes(x = pre_cov_adm_rate))

# hypo_all_trend_hb_ans %>% 
#   summarise(cov_sd = sd(cov_adm_rate),
#          pre_cov_sd = sd(pre_cov_adm_rate))+

```
```{r}
# boxplot
hypo_cov <- hypo_all_trend_hb_ans %>% 
  select(-pre_cov_adm_rate, -pre_cov_adm) %>% 
  mutate(data_type = "cov") %>% 
  relocate(data_type, .before = year) %>% 
  rename(rate = cov_adm_rate,
         adm_num = cov_adm)

hypo_pre_cov <- hypo_all_trend_hb_ans %>% 
  select(-cov_adm_rate, -cov_adm)%>% 
  mutate(data_type = "pre_cov") %>% 
  relocate(data_type, .before = year)%>% 
  rename(rate = pre_cov_adm_rate,
         adm_num = pre_cov_adm)

hypo_all = bind_rows(hypo_cov,hypo_pre_cov) %>% 
  group_by(data_type, year, quarter, week_ending) %>% 
  summarise(adm_num = sum(adm_num))

hypo_all %>% 
  ggplot()+
  aes(x = adm_num, y = data_type, group = data_type)+
  geom_boxplot(show.legend = TRUE)+
  stat_summary(fun.y = "mean", geom = "point", colour = "red")+
    labs(title = "Covid Rate Boxplot")

```



# Hypothesis test - Covid Impact Hospital Admission

#### Hypothesis Test Setup
* Test type: 2 Sample paired Test

* Dataset: Hospital admission Number between Covid Data(2020-2021) and Pre-Covid Data(2018-2019)

* $\alpha = 0.05$

* $\Delta\mu = \mu_{cov} - \mu_{pre cov}$

* $H_{0}: \Delta\mu = 0$

* $H_{\alpha}: \Delta\mu > 0$


```{r}
hypo_covid <- hypo_all_trend_hb_ans %>% 
  select(-pre_cov_adm_rate, - cov_adm_rate) %>% 
  group_by(year,quarter,week_ending) %>% 
  summarise(cov_adm = sum(cov_adm),
            pre_cov_adm = sum(pre_cov_adm)) %>% 
    mutate(diff_adm = cov_adm - pre_cov_adm)


hypo_covid %>% 
  ggplot()+
  aes(x = diff_adm)+
  geom_boxplot()

```


```{r}
null_distribution <- hypo_covid %>% 
  specify(response = diff_adm) %>% 
  hypothesize(null = "point", mu = 0) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

```{r}
observed_stat <- hypo_covid %>% 
  specify(response = diff_adm) %>% 
  calculate(stat = "mean")
observed_stat
```
```{r}
null_distribution %>%
  visualise() +
  shade_p_value(obs_stat = observed_stat, direction = "right")
```
```{r}
p_value <- null_distribution %>% 
  get_pvalue(obs_stat = observed_stat,
             direction = "right")
p_value
```


```{r}
df_ha_ans <- read_csv("clean_data/df_ha_ans.csv")

ha_ans <- df_ha_ans %>% 
  filter(admission_type == c("All Inpatients and Day cases", "Emergency Inpatients", "Not Specified")) %>% 
  group_by(year, quarter) %>% 
  rename(adm_num = episodes) %>% 
  summarise(adm_num = sum(adm_num)) %>% 
  mutate(quarter = factor(quarter))

```

# Hypothesis Test Summary - Covid Impact Hospital Admission 

`The p-value of ~1 is above α=0.05, so we reject the null hypothesis and conclude that, there is no significant evidence tha the admission is higher during covid period than pre-covid period.`


#########

# Hypothesis Test  - Winter Impact 

# Hypothesis Setup  - Winter Impact(Q1) on Spring (Q2) 
# H0: mu_winter - mu_spring = 0
# Ha: mu_winter - mu_spring > 0
```{r}
hypo_season_spring <- ha_ans %>% 
  filter(quarter == "2" | quarter == "1")
```
```{r}
null_distribution <- hypo_season_spring %>% 
  specify(adm_num ~ quarter) %>% 
  hypothesize(null = "independence") %>% 
  generate(reps = 5000, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("1", "2")) 
```

```{r}
observed_stat <- hypo_season_spring %>% 
  specify(adm_num ~ quarter) %>% 
  calculate(stat = "diff in means", order = c("1", "2")) 
  
observed_stat
```

```{r}
null_distribution %>%
  visualise() +
  shade_p_value(obs_stat = observed_stat, direction = "right")
```

```{r}
p_value <- null_distribution %>%
  get_p_value(obs_stat = observed_stat, direction = "right")

p_value
```
# Winter compare to Summer

`The p-value of 0.1676 is above α=0.05, so we failed to reject the null hypothesis and conclude that, there is no significant evidence that the admission during winter period is higher than spring`




# ############################
# Hypothesis Setup  - Winter Impact(Q1) on Summer (Q2) 
# H0: mu_winter - mu_spring = 0
# Ha: mu_winter - mu_spring > 0
```{r}
hypo_season_summer <- ha_ans %>% 
  filter(quarter == "3" | quarter == "1")
```
```{r}
null_distribution <- hypo_season_summer %>% 
  specify(adm_num ~ quarter) %>% 
  hypothesize(null = "independence") %>% 
  generate(reps = 5000, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("1", "3")) 
```

```{r}
observed_stat <- hypo_season_summer %>% 
  specify(adm_num ~ quarter) %>% 
  calculate(stat = "diff in means", order = c("1", "3")) 
  
observed_stat
```

```{r}
null_distribution %>%
  visualise() +
  shade_p_value(obs_stat = observed_stat, direction = "right")
```

```{r}
p_value <- null_distribution %>%
  get_p_value(obs_stat = observed_stat, direction = "right")

p_value
```



# ######
# ############################
# Hypothesis Setup  - Winter Impact(Q1) on Summer (Q2) 
# H0: mu_winter - mu_spring = 0
# Ha: mu_winter - mu_spring > 0
```{r}
hypo_season_autumn <- ha_ans %>% 
  filter(quarter == "4" | quarter == "1")
```
```{r}
null_distribution <- hypo_season_autumn %>% 
  specify(adm_num ~ quarter) %>% 
  hypothesize(null = "independence") %>% 
  generate(reps = 5000, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("1", "4")) 
```

```{r}
observed_stat <- hypo_season_autumn %>% 
  specify(adm_num ~ quarter) %>% 
  calculate(stat = "diff in means", order = c("1", "4")) 
  
observed_stat
```

```{r}
null_distribution %>%
  visualise() +
  shade_p_value(obs_stat = observed_stat, direction = "right")
```

```{r}
p_value <- null_distribution %>%
  get_p_value(obs_stat = observed_stat, direction = "right")

p_value
```



# DUMP
```{r}

# hypo_season_all <- 
# 
# pre_cov_hypo_season <- hypo_season_all %>% 
#   select(-cov_adm) %>%
#   rename(adm_num = pre_cov_adm) %>% 
#   mutate(data_type = "pre_cov") %>% 
#   relocate(data_type, .before = year)
# 
# cov_hypo_season <- hypo_season_all %>% 
#   select(-pre_cov_adm)%>%
#     rename(adm_num = cov_adm) %>% 
#   mutate(data_type = "cov") %>% 
#   relocate(data_type, .before = year)
# 
# hypo_season = bind_rows(cov_hypo_season, pre_cov_hypo_season) %>% 
#   group_by(quarter, week_ending) %>% 
#   summarise(adm_num = sum(adm_num)) %>% 
#   mutate(quarter = factor(quarter))
# 
# 
# hypo_season %>% 
#   ggplot()+
#   aes(x = adm_num, y = quarter, group = quarter)+
#   geom_boxplot()

```
```

