---
title: "R Notebook"
output: html_notebook
---

```{r libraries, echo=FALSE, include=FALSE}
library(janitor)
library(tidyverse)
library(ggplot2)
library(here)
library(leaflet)
library(stringr)
```

```{r}
here::here()
```

```{r data}
df_ha_dep <-read_csv(here("../phs_rshiny_dashboard_grp1/clean_data/df_ha_dep.csv"))

df_hospital_location <-read_csv(here("../phs_rshiny_dashboard_grp1/clean_data//df_hospital_location.csv"))

df_ha_dep <-read_csv(here("../phs_rshiny_dashboard_grp1/clean_data/df_ha_dep.csv"))

df_cov_hb <- read_csv(here("../phs_rshiny_dashboard_grp1/clean_data/df_cov_hb_dep.csv"))

```

Deprivation breakdown for inpatient and daycase activity (episode and stays) by health board of treatment of the patients.

## Definition of SIMD

Scottish Index of Multiple Deprivation (SIMD) quintile; 1(Most Deprived) - 5(Least Deprived); Most appropriate SIMD release used for each year

## Dataset descriptions:

df_ane_hb_dep: 2020-1, weekly data, health board, attendances

df_cov_hb_dep: 2020-1, covid, weekly data, health board, attendances, attendance type

df_cov_hscp_dep: 2020-1, covid, weekly data, hscp, attendances, attendance type

df_ha_dep: 2016-2021, quarterly, hospital level, attendances, attendance type

# Hospital Admission by Deprivation ----
```{r}
# Remove missing simd values
df_ha_dep <- df_ha_dep %>%
  filter(!is.na(simd))
```

```{r}
# add in a quarter and year variable
df_ha_dep <- df_ha_dep %>%
  mutate(q_year = str_c(year, " Q",quarter))
```

```{r}
# Pivot wider to add simd as a column
df_ha_dep_wide <- df_ha_dep %>%
  pivot_wider(names_from = simd, values_from = episodes)
```

```{r}
# rename columns
df_ha_dep_wide <- df_ha_dep_wide %>%
    `colnames<-`(c("year", "quarter", "hb", "location", "admission_type",
                   "length_of_episode", "average_length_of_episode", "stays",
                   "length_of_stay", "average_length_of_stay","q_year", "simd_1",
                   "simd_2", "simd_3", "simd_4", "simd_5"))
```

```{r}
df_ha_dep_summary <- df_ha_dep %>%
  filter(admission_type == "Emergency Inpatients") %>%
  filter(length_of_stay != 0) %>%
  filter(length_of_episode != 0) %>%
  group_by(hb, simd)
```

```{r}
df_ha_dep_summary %>%
  ggplot(aes(x = q_year, fill = hb)) +
  geom_bar(position = "dodge", colour="white", alpha=0.3) +
  facet_wrap(~hb) +
  labs(
    x = "\nYear",
    y = "Count\n",
    title = "Deprivation",
    subtitle = "\n"
    )
```



```{r}
df_ha_dep %>%
  group_by(admission_type) %>%
  summarise(count = n())
```




```{r}
df_ha_dep_wide %>%
  ggplot(aes(x = year, fill = hb)) +
  geom_bar(position = "dodge", colour="white", alpha=0.3) +
  facet_wrap(~hb) +
  labs(
    x = "\nYear",
    y = "Count\n",
    title = "Deprivation",
    subtitle = "\n"
    )
```

```{r}
df_ha_dep %>%
  mutate(simd = factor(simd)) %>%
  ggplot(aes(x = year, fill = simd)) +
  geom_bar(position = "dodge", colour="white", alpha=0.3) +
  facet_wrap(~simd) +
  labs(
    x = "\nYear",
    y = "Count\n",
    title = "Deprivation",
    subtitle = "\n"
    )
```

```{r}
df_ha_dep %>%
  mutate(simd = factor(simd)) %>%
  group_by(simd) %>%
  ggplot(aes(x = length_of_stay, y = length_of_episode, group = simd, colour = simd)) +
  geom_point() +
 # facet_wrap(~simd) +
  labs(
    x = "\nYear",
    y = "Count\n",
    title = "Deprivation",
    subtitle = "\n"
    )



```
